---
output:
  github_document:
    html_preview: false
---

# prais

[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/prais)](https://cran.r-project.org/package=prais)
[![R-CMD-check](https://github.com/FranzMohr/prais/workflows/R-CMD-check/badge.svg)](https://github.com/FranzMohr/prais/actions)

## Overview

prais implements the Prais-Winsten estimator for models with strictly exogenous regressors and AR(1) serial correlation of the errors.

## Installation

### CRAN 

```{r cran, eval = FALSE}
install.packages("prais")
```

### Development version

```{r github, eval = FALSE}
# install.packages("devtools")
devtools::install_github("franzmohr/prais")
```

## Usage

```{r usage, message = FALSE}
# Load the package
library(prais)

# Load the data
# install.packages("wooldridge")
library(wooldridge)
data("barium")

pw <- prais_winsten(lchnimp ~ lchempi + lgas + lrtwex + befile6 + affile6 + afdec6,
                    data = barium, index = "t")
summary(pw)
```

## Robust standard errors

### White's estimator

```{r, message = FALSE}
library(lmtest)

coeftest(pw, vcov. = vcovHC(pw, "HC1"))
```

### Panel-corrected standard errors (PCSE)

Estimate a panel model, for which PCSE should be obtained.

```{r, message = FALSE}
# Example 2 in the documentation of Stata function xtpcse

# Load data
data <- haven::read_dta("http://www.stata-press.com/data/r14/grunfeld.dta")

# Estimate
x <- prais_winsten(invest ~ mvalue + kstock, data = data, index = c("company", "year"),
                   twostep = TRUE, panelwise = TRUE, rhoweight = "T1")

# Results
summary(x)
```
Obtain PCSE by using only those residuals from periods that are common to all panels by setting `pairwise = FALSE`.

```{r}
coeftest(x, vcov. = vcovPC(x, pairwise = FALSE))
```

Obtain PCSE by using all observations that can be matched by period between two panels by setting `pairwise = TRUE`.

```{r}
coeftest(x, vcov. = vcovPC(x, pairwise = TRUE))
```

## References

Beck, N. L. and Katz, J. N. (1995): What to do (and not to do) with time-series cross-section data. American Political Science Review 89, 634-647.

Prais, S. J. and Winsten, C. B. (1954): Trend Estimators and Serial Correlation. Cowles Commission Discussion Paper, 383 (Chicago).

Wooldridge, J. M. (2016). Introductory Econometrics. A Modern Approach. 6th ed. Mason, OH: South-Western Cengage Learning.
