---
output:
  github_document:
    html_preview: false
---

# prais

[![CRAN status](https://www.r-pkg.org/badges/version/prais)](https://cran.r-project.org/package=prais)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/prais)](https://cran.r-project.org/package=prais)
[![Total downloads](https://cranlogs.r-pkg.org/badges/grand-total/prais)](https://cran.r-project.org/package=prais)
[![R-CMD-check](https://github.com/franzmohr/prais/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/franzmohr/prais/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/franzmohr/prais/graph/badge.svg)](https://app.codecov.io/gh/franzmohr/prais)
[![License: GPL (>= 2)](https://img.shields.io/badge/license-GPL%20%28%3E%3D%202%29-blue.svg)](https://www.gnu.org/licenses/gpl-2.0)
[![DOI](https://img.shields.io/badge/DOI-10.5281%2Fzenodo.22837936-blue.svg)](https://doi.org/10.5281/zenodo.22837936)

[![GitHub Sponsors](https://img.shields.io/badge/Sponsor-%E2%9D%A4-ea4aaa?logo=github-sponsors&logoColor=white)](https://github.com/sponsors/franzmohr)
[![Buy Me a Coffee](https://img.shields.io/badge/Buy%20Me%20a%20Coffee-FFDD00?logo=buymeacoffee&logoColor=black)](https://buymeacoffee.com/franzmohr)

## Overview

`prais` implements the Prais-Winsten estimator for models with strictly exogenous regressors and AR(1) serial correlation of the errors.

## Installation

### CRAN 

```{r cran, eval = FALSE}
install.packages("prais")
```

### Development version

```{r github, eval = FALSE}
# install.packages("devtools")
devtools::install_github("franzmohr/prais")
```

## Usage

```{r usage, message = FALSE}
# Load the package
library(prais)

# Load the data
data("barium")

pw <- prais_winsten(lchnimp ~ lchempi + lgas + lrtwex + befile6 + affile6 + afdec6,
                    data = barium, index = "t")
summary(pw)
```

### Panel models estimated with plm

A pooled panel model that was estimated with `plm::plm` can be passed to
`prais_winsten` instead of the formula, the data and the index, which are taken
from the model. Since the Prais-Winsten estimator obtains all its estimates by
ordinary least squares, only models that were estimated with `model = "pooling"`
are supported. Fixed effects can be estimated by adding the dummy variables to
the formula, as in `y ~ x + factor(id)`.

```{r plm, message = FALSE, eval = FALSE}
library(plm)

data("Grunfeld", package = "plm")

pooled <- plm(inv ~ value + capital, data = Grunfeld,
              index = c("firm", "year"), model = "pooling")

summary(prais_winsten(pooled, panelwise = TRUE, rhoweight = "T1"))
```

## Confidence intervals and predictions

`confint` gives confidence intervals for the coefficients. They are based on the
*t* distribution with the residual degrees of freedom of the model, so they agree
with the p-values that `summary` reports.

```{r confint}
confint(pw)
```

`vcov` returns the covariance matrix of the coefficients those standard errors
come from. Functions that obtain the covariance matrix from a model, such as
`lmtest::coeftest`, therefore work without being passed a matrix explicitly.

```{r vcov}
sqrt(diag(vcov(pw)))
```

`predict` returns the conditional mean of the model. Its standard errors and
confidence intervals describe that mean and come from the same covariance matrix,
so they agree with `confint`. The AR(1) structure of the errors is not used, which
is why intervals for an individual observation are not available.

```{r predict-interval}
head(predict(pw, interval = "confidence"))
```

With `se.fit = TRUE` the standard errors are returned alongside the predictions,
together with the residual degrees of freedom and the residual standard error.

```{r predict-se}
fcst <- predict(pw, se.fit = TRUE)

head(fcst$se.fit)
```

Argument `newdata` predicts for observations that were not part of the estimation.
The variables of the formula do not have to be transformed beforehand.

## Robust standard errors

### White's estimator

```{r, message = FALSE}
library(lmtest)

coeftest(pw, vcov. = vcovHC(pw, "HC1"))
```

### Panel-corrected standard errors (PCSE)

Estimate a panel model, for which PCSE should be obtained.

```{r, message = FALSE}
# Example 2 in the documentation of Stata function xtpcse

# Load data
data <- haven::read_dta("http://www.stata-press.com/data/r14/grunfeld.dta")

# Estimate
x <- prais_winsten(invest ~ mvalue + kstock, data = data, index = c("company", "year"),
                   twostep = TRUE, panelwise = TRUE, rhoweight = "T1")

# Results
summary(x)
```
Obtain PCSE by using only those residuals from periods that are common to all panels by setting `pairwise = FALSE`.

```{r}
coeftest(x, vcov. = vcovPC(x, pairwise = FALSE))
```

Obtain PCSE by using all observations that can be matched by period between two panels by setting `pairwise = TRUE`.

```{r}
coeftest(x, vcov. = vcovPC(x, pairwise = TRUE))
```

## Tidy model output

Methods for the generics `tidy`, `glance` and `augment` of package
[broom](https://cran.r-project.org/package=broom) summarise the coefficients, the
goodness of fit and the fitted values in tidy data frames. They are the interface
through which packages for publishable regression output, such as `gtsummary` and
`modelsummary`, obtain the results of a model. `broom` has to be installed to use
them.

```{r broom, message = FALSE}
library(broom)

tidy(pw, conf.int = TRUE)
```

```{r glance}
glance(pw)
```

`glance` also reports the AR(1) coefficient and the Durbin-Watson statistics of
the original and the transformed model. All three are `NA` if the model was
estimated with `panelwise = TRUE`, because there is one coefficient per panel in
that case.

```{r augment}
augment(pw)
```

## Citation

To cite `prais` in publications, use

```{r citation, eval = FALSE}
citation("prais")
```

Releases are archived on Zenodo. To cite the package as a whole, rather than one
particular version, use the concept DOI
[10.5281/zenodo.22837936](https://doi.org/10.5281/zenodo.22837936), which always
resolves to the most recent release.

## References

Beck, N. L. and Katz, J. N. (1995): What to do (and not to do) with time-series cross-section data. American Political Science Review 89, 634-647.

Prais, S. J. and Winsten, C. B. (1954): Trend Estimators and Serial Correlation. Cowles Commission Discussion Paper, 383 (Chicago).

Wooldridge, J. M. (2016). Introductory Econometrics. A Modern Approach. 6th ed. Mason, OH: South-Western Cengage Learning.
