---
title: "Estimating models with AR(1) errors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimating models with AR(1) errors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## When to use the estimator

If the errors of a linear regression follow an AR(1) process, the coefficients of
ordinary least squares are still unbiased, but they are no longer efficient and
their standard errors are wrong. The Prais-Winsten estimator corrects for this by
transforming the data so that the transformed errors are serially uncorrelated,
and by estimating the coefficients and the autocorrelation of the errors in turn
until the latter converges.

The estimator assumes that the regressors are *strictly exogenous*. A model that
contains a lagged dependent variable does not satisfy this, and its estimates
would be inconsistent. The transformation also assumes an AR(1) process, not a
higher order one.

The Cochrane-Orcutt estimator solves the same problem by dropping the first
observation of the sample. Prais-Winsten keeps it and rescales it instead, which
matters in short samples.

```{r}
library(prais)
```

## A simple time series model

The package ships the `barium` data, monthly observations from the U.S. chemical
industry. The column `t` is the time index.

```{r}
data("barium")

str(barium[, c("lchnimp", "lchempi", "lgas", "lrtwex", "t")])
```

The model is specified as it would be for `lm`. Argument `index` names the time
variable, by which the observations are ordered before the transformation is
applied. Without it the estimator would have to assume that the rows already are
in the right order.

```{r}
pw <- prais_winsten(lchnimp ~ lchempi + lgas + lrtwex + befile6 + affile6 + afdec6,
                    data = barium, index = "t")
```

The iterations are reported as they happen, which shows how the estimate of the
AR(1) coefficient *rho* settles down. They are ordinary messages and can be
silenced with `suppressMessages`.

### Reading the output

```{r}
summary(pw)
```

Three parts of the output are specific to this estimator.

* The **AR(1) coefficient** and the number of iterations it took to converge. A
value clearly different from zero is what justifies using the estimator in the
first place.
* The **residuals of the transformed model**, rather than those of the original
one. The transformed model is what is actually fitted by least squares, so its
residuals are the ones whose properties matter.
* Two **Durbin-Watson statistics**. The first belongs to the original model and
the second to the transformed one. A value near 2 indicates no first order serial
correlation, so the pair shows whether the transformation did its job.

It is worth comparing the result with ordinary least squares.

```{r}
ols <- lm(lchnimp ~ lchempi + lgas + lrtwex + befile6 + affile6 + afdec6,
          data = barium)

round(cbind("OLS" = coef(ols), "PW" = coef(pw),
            "SE (OLS)" = sqrt(diag(vcov(ols))),
            "SE (PW)" = sqrt(diag(vcov(pw)))), 4)
```

Some coefficients move a good deal, `lgas` most of all. Every standard error is
larger than its counterpart from least squares, here by between eight and thirty
per cent, because least squares treats positively correlated errors as though
they carried more independent information than they do. Which of the two effects
dominates depends on the data; the point is that the corrected standard errors
are the ones to report.

### Confidence intervals and predictions

`confint` uses the *t* distribution with the residual degrees of freedom of the
model, so the intervals agree with the p-values of the summary.

```{r}
confint(pw)
```

`predict` returns the conditional mean of the model. With
`interval = "confidence"` the interval of that mean is added. The AR(1) structure
of the errors is not used for the prediction, which is why an interval for an
individual future observation is not available.

```{r}
head(predict(pw, interval = "confidence"))
```

Passing `newdata` predicts for observations that were not part of the estimation.
The variables of the formula do not have to be transformed beforehand.

### Robust standard errors

If the errors are heteroskedastic on top of being serially correlated, the
covariance matrix of the transformed model can be made robust with `vcovHC`.

```{r}
sqrt(diag(vcovHC(pw, type = "HC1")))
```

## A panel model

For panel data, `index` takes two names, the ID variable first and the time
variable second. The transformation is then applied within each panel, so that
the last observation of one panel is never differenced against the first
observation of the next.

A small balanced panel with a known AR(1) coefficient makes it easy to see what
the estimator recovers.

```{r}
set.seed(123)

n_id <- 8
n_time <- 25
rho <- .6

panel <- do.call(rbind, lapply(seq_len(n_id), function(i) {
  u <- rnorm(n_time, 0, 2)
  for (j in 2:n_time) {
    u[j] <- u[j] + rho * u[j - 1]
  }
  x <- rnorm(n_time, 10, 3)
  data.frame(id = i, year = seq_len(n_time), x = x, y = 5 + 2 * x + u)
}))

head(panel)
```

```{r}
pw_panel <- prais_winsten(y ~ x, data = panel, index = c("id", "year"))

summary(pw_panel)
```

The coefficient on `x` is recovered almost exactly, and the Durbin-Watson
statistic of the transformed model is much closer to 2 than that of the original
one. The estimate of *rho* lands below the 0.6 the errors were generated with.
That is the usual downward bias of an autocorrelation estimated from the
residuals of a short series, which is of the order of one divided by the number
of periods per panel, and it shrinks as the panels get longer. The correction
does not depend on hitting *rho* exactly.

### One coefficient per panel

By default a single *rho* is estimated from all panels together. With
`panelwise = TRUE` every panel gets its own, which is appropriate if the panels
differ in how persistent their errors are.

```{r}
pw_pw <- prais_winsten(y ~ x, data = panel, index = c("id", "year"),
                       panelwise = TRUE)

round(pw_pw$rho[nrow(pw_pw$rho), ], 3)
```

Panel specific estimates come from short series and are therefore noisy. Argument
`rhoweight` averages them into a single coefficient, weighted by the length of
each panel, which trades some flexibility for a more stable estimate.

```{r}
pw_t1 <- prais_winsten(y ~ x, data = panel, index = c("id", "year"),
                       panelwise = TRUE, rhoweight = "T1")

pw_t1$rho[nrow(pw_t1$rho), ]
```

### Panel-corrected standard errors

Errors are often correlated across panels within the same period. `vcovPC`
implements the panel-corrected standard errors of Beck and Katz (1995) for that
case.

```{r}
sqrt(diag(vcovPC(pw_panel)))
```

`pairwise = FALSE`, the default, uses only the periods that all panels have in
common. `pairwise = TRUE` uses every period that a given pair of panels shares,
which matters for unbalanced panels.

### Gaps in the time variable

Two observations that lie *k* periods apart are correlated by *rho^k* under an
AR(1) process. The transformation accounts for this, so a panel with missing
periods is not treated as if its observations were consecutive.

```{r}
gapped <- panel[panel$year != 10, ]

pw_gap <- prais_winsten(y ~ x, data = gapped, index = c("id", "year"))

pw_gap$rho[nrow(pw_gap$rho), ]
```

The periods are counted in steps of the greatest common divisor of the
differences of the time variable, so quarterly, monthly or biennial data need no
further arguments, and `Date` variables are supported.

### Models estimated with plm

A pooled panel model that was fitted with `plm::plm` can be handed to
`prais_winsten` in place of the formula, the data and the index, which are then
taken from the model.

```{r, eval = requireNamespace("plm", quietly = TRUE)}
data("Grunfeld", package = "plm")

pooled <- plm::plm(inv ~ value + capital, data = Grunfeld,
                   index = c("firm", "year"), model = "pooling")

coef(suppressMessages(prais_winsten(pooled)))
```

Only models estimated with `model = "pooling"` are supported, because the
Prais-Winsten estimator obtains all of its estimates by ordinary least squares.
The within and the random effects estimator apply a transformation of their own,
which does not leave the AR(1) structure of the errors intact. Fixed effects can
be obtained by putting the dummy variables into the formula, as in
`y ~ x + factor(id)`.

## Tidy output

Methods for the generics of
[broom](https://cran.r-project.org/package=broom) are available, which packages
for publishable regression tables use to read a model.

```{r, eval = requireNamespace("broom", quietly = TRUE)}
broom::tidy(pw_panel, conf.int = TRUE)
```

## References

Beck, N. L. and Katz, J. N. (1995): What to do (and not to do) with time-series
cross-section data. *American Political Science Review* 89, 634-647.

Prais, S. J. and Winsten, C. B. (1954): Trend Estimators and Serial Correlation.
*Cowles Commission Discussion Paper*, 383 (Chicago).

Wooldridge, J. M. (2016): *Introductory Econometrics. A Modern Approach*. 6th ed.
Mason, OH: South-Western Cengage Learning.
